---
title: "**Data Science Career Paths & Skillsets in 2021**"
subtitle: "*A glimpse of Kaggle’s State of Data Science and Machine Learning Survey*"
author:
  - Jingsong Gao, jg2109@georgetown.edu
  - Ercong Luo, el890@georgetown.edu
  - Rui Qiu, rq47@georgetown.edu
date: "`r Sys.Date()`"
output:
  rmdformats::robobook
  # rmdformats::readthedown
---

```{r setup, include=FALSE}
# invalidate cache when the package version changes
knitr::opts_chunk$set(cache = TRUE,
                      echo = FALSE,
                      eval = TRUE,
                      tidy = FALSE,
                      warning = FALSE,
                      cache.extra = packageVersion('tidyverse'))
options(htmltools.dir.version = FALSE)

if (!require("pacman")) install.packages("pacman")
pacman::p_load(tidyverse, ggthemes, latex2exp, glue,
               hrbrthemes, plotly, stringr, DT, extrafont,
               tidymodels, rmdformats)

set.seed(511)

ggplot2::theme_set(
    theme_fivethirtyeight() +
    theme(
        text = element_text(family = "Roboto Condensed"),
        title = element_text(size = 14),
        plot.subtitle = element_text(size = 12),
        plot.caption = element_text(size = 10),
        axis.title = element_text(size = 14),
        axis.text = element_text(size = 12),
        panel.grid.minor.x = element_blank()
    )
)
```

# Introduction

Address the data science problem

# Data Cleaning

```{r pkg, message=FALSE}
col_names <- names(read_csv(
  "data/kaggle_survey_2021_responses.csv",
  n_max=0))
dat <- read_csv(
  "data/kaggle_survey_2021_responses.csv",
  col_names = col_names, skip=2)

dat <- dat %>%
  filter(Q3=="United States of America" )

job.dat <- dat %>% 
    filter(Q5 %in% c("Data Analyst",
                     "Data Engineer",
                     "Data Scientist",
                     "Machine Learning Engineer",
                     "Software Engineer",
                     "Statistician",
                     "Student")) %>%
    mutate(Q25 = str_remove_all(Q25, "[$,]")) %>%
    mutate(Q25 = str_replace(Q25, ">1000000", "1000000-2000000")) %>%
    separate(Q25, into = c("salary_lb", "salary_ub"), sep = "-") %>%
    mutate(salary_lb = as.numeric(salary_lb)) %>%
    mutate(salary_ub = as.numeric(salary_ub))
```

# The Analysis and Results

- [ ] List of questions.

Since the one large data science problem can be divided into several smaller bits, we decide to combine the analysis and the interpretation sections together. And the questions will be addressed and discussed one by one.

The following questions are about data science career paths.

## Job Titles Distributions

What percentage of the survey respondents are working under these job titles? 

```{r q1-job}
# Q1

jtitle <- sort(table(dat$Q5), decreasing = T) %>% 
    as.data.frame() %>%
    as.tibble()
jtitle <- rename(jtitle, `Job Title` = Var1)

ggplot(jtitle, aes(x="", y=Freq, fill=`Job Title`)) +
    geom_bar(stat="identity", width=1, color="white") +
    coord_polar("y", start=0) + 
    theme_void() 
```

## Salaries

What are the statistics on salaries for these job titles?

The US minimum wage is 7.25 per hour, multiply that for a total $260$ typical number of workdays a year:

$$7.25 \frac{\$}{\text{hour}} \cdot 8 \frac{\text{hours}}{\text{day}} \cdot 260\frac{\text{days}}{\text{year}} = 15080.$$

The new categories* for salaries will be:
- poverty: below the federal minimum wage
- low: 15,000 to 49,999
- medium: 50,000 to 99,999
- high: 100,000 to 199,999
- very high: 200,000 to 499,999
- highest: >= 500,000

*loosely based on US federal income tax brackets. 


```{r q2}
# Q2

# To change salary categories into FACTOR dtype with descending labels
poverty   = c("$0-999", "1,000-1,999" , "2,000-2,999", "3,000-3,999", "4,000-4,999", "5,000-7,499", "7,500-9,999",
             "10,000-14,999") 
low       = c("15,000-19,999", "20,000-24,999", "25,000-29,999", "30,000-39,999", "40,000-49,999")
medium    = c("50,000-59,999", "60,000-69,999", "70,000-79,999", "80,000-89,999", "90,000-99,999") 
high      = c("100,000-124,999", "125,000-149,999", "150,000-199,999")
very_high = c("200,000-249,999", "250,000-299,999", "300,000-499,999")
highest   = c("$500,000-999,999", ">$1,000,000")

dat$Q25[dat$Q25 %in% poverty] <- "poverty"
dat$Q25[dat$Q25 %in% low] <- "low"
dat$Q25[dat$Q25 %in% medium] <- "medium"
dat$Q25[dat$Q25 %in% high] <- "high"
dat$Q25[dat$Q25 %in% very_high] <- "very high"
dat$Q25[dat$Q25 %in% highest] <- "highest"
dat$Q25 = factor(dat$Q25, levels = c("poverty", "low", "medium", "high", "very high", "highest"), ordered = T)

data_side <- c("Data Scientist", "Data Analyst", "Business Analyst", "Data Engineer", "Statistician", "DBA/Database Engineer")
swe_side <- c("Software Engineer", "Machine Learning Engineer", "Program/Project Manager", "Product Manager") 
academia <- c("Student", "Other", "Research Scientist") 

dat[dat$Q5 %in% data_side & !is.na(dat$Q5) & !is.na(dat$Q25), ] %>%
  ggplot( aes(x=Q5, y=Q25, color = Q5)) +
    geom_count() + 
    ggtitle("Two-Way Salary Visualizations: Data-Oriented Jobs") +
    xlab("") +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))


dat[dat$Q5 %in% swe_side & !is.na(dat$Q5) & !is.na(dat$Q25), ] %>%
  ggplot( aes(x=Q5, y=Q25, color = Q5)) +
    geom_count() + 
    ggtitle("Two-Way Salary Visualizations: Engineering-Oriented Jobs") +
    xlab("") + 
    theme(axis.text.x = element_text(angle = 45, hjust = 1))

dat[dat$Q5 %in% academia & !is.na(dat$Q5) & !is.na(dat$Q25), ] %>%
  ggplot( aes(x=Q5, y=Q25, color = Q5)) +
    geom_count() + 
    ggtitle("Two-Way Salary Visualizations: Academic Jobs and Others") +
    xlab("") + 
    theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

A chi-squared test of independence can be used to test for the statistical significance in the difference in categorical frequencies in each salary category for each job title. On the one hand, the test is done on data-oriented data science roles, see below: 

```{r q2-a}
expectedcounts <- function(A){
  r <- rowSums(A)
  c <- colSums(A)
  N = sum(A)
  expected <- outer(r,c)/N
  return(expected)
}


q2_table.1 = table(dat$Q25[dat$Q5 %in% data_side], dat$Q5[dat$Q5 %in% data_side])
q2_table.df = q2_table.1 %>% as.data.frame() %>% as.tibble()
q2_expectedcounts <- expectedcounts(q2_table.1) %>% as.data.frame() %>% as.tibble()

ggplot(q2_table.df, aes(x = Var2, y = Freq, fill = Var1)) +
    geom_col(position = "dodge") +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))

(chisq.test(q2_table.1, simulate.p.value = T, B = 10000))
```

Qualitatively, it can be seen that each job title has a seemingly different categorical distribution over the salary brackets. The chi-squared test using simulated p-values reveals that this difference is statistically significant as well with a p-value of $0.0004998$. At an alpha-level of 0.05, the null hypothesis that the frequency of counts over salary brackets is not independent of the job title. 

Because the original salary data is categorical instead of a numerical estimation, a two-sample t-test to test the difference of mean salary is not possible. But the visualization above reveals that for data scientists and data engineers, there are more people in medium and high salary brackets than low salary brackets; the opposite is true for business analyst and data analyst. The visualization alone is able to reveal that "analyst" is considered a more junior role compared to an "engineer" or a "data scientist". There are just too few DBA/Database engineers and statisticians represented in the dataset for a conclusion to be drawn for these categories.

It is also worth noting that most of the highest earning ($>\$500,000$) people in the dataset are data scientists. 

Below, the analysis focuses on the job titles that have a focus on software and product management, such as machine learning engineer (MLE), software engineer (SWE), and product manager (PM). 

```{r q2-b}
q2_table.2 = table(dat$Q25[dat$Q5 %in% swe_side], dat$Q5[dat$Q5 %in% swe_side])
q2_table.df2 = q2_table.2 %>% as.data.frame() %>% as.tibble()
q2_expectedcounts2 <- expectedcounts(q2_table.2) %>% as.data.frame() %>% as.tibble()

ggplot(q2_table.df2, aes(x = Var2, y = Freq, fill = Var1)) +
    geom_col(position = "dodge") +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))

(chisq.test(q2_table.2, simulate.p.value = T, B = 10000))
```

As can be seen above, more MLEs command the highest category of salary than any other job title. The average MLE will earn at or above the medium salary bracket. As for PMs, very few of them ear a low salary; in fact, most of them are in the medium to very high brackets of income. The salary distribution for project managers is similar to that of PMs, except there are significantly more professionals in the low income bracket. Last but not least, most SWEs command a medium to high salary, with still a significant number of professional earning outside this range on both the high end and the lower end. 

## Requried Education Levels

What levels of education are required for these job titles?

### Aggregate Level

```{r q3}
# Q3

edu <- sort(table(dat$Q4), decreasing = T) %>% 
    as.data.frame() %>%
    as.tibble()
edu <- rename(edu, `Degree` = Var1)

ggplot(edu, aes(x="", y=Freq, fill=`Degree`)) +
    geom_bar(stat="identity", width=1, color="white") +
    coord_polar("y", start=0) + 
    theme_void() 
```

```{r q3-a}
Q3_temp <- sort(table(dat$Q4), decreasing = T)
(Q3_temp <- Q3_temp / sum(Q3_temp))
```


At an aggregate level, it is evident that more than half of the data science community on Kaggle has a degree that is above the undergraduate level. Combining percentages of people with a master's degree or a doctoral degree (including a professional doctoral degree) gives us that $64.7/%$ of respondents hold a graduate degree. For the combination of programming skill level requirements, as well as the mount of requisite knowledge in statistics machine learning, this is an expected phenomenon. 

### Most professional data scientists have a master's degree or above:

```{r q3-b}
edu.1 <- sort(table(dat[dat$Q5 == "Data Scientist",]$Q4), decreasing = T) %>% 
    as.data.frame() %>%
    as.tibble()
edu.1 <- rename(edu.1, `Degree` = Var1)

ggplot(edu.1, aes(x="", y=Freq, fill=`Degree`)) +
    geom_bar(stat="identity", width=1, color="white") +
    coord_polar("y", start=0) + 
    theme_void() 
```

### The same doesn't necessary go for software engineering

```{r q3-c}
EDU <- table(dat$Q5, dat$Q4) %>% 
    as.data.frame() %>%
    as.tibble()
EDU <- rename(EDU, `Job Title` = Var1)
EDU <- rename(EDU, `Degree` = Var2)

# ggplot(EDU, aes(x="", y=Freq, fill=`Degree`)) +
#     geom_bar(stat="identity", width=1, color="white") +
#     coord_polar("y", start=0) + 
#     theme_void() + 
#     facet_wrap(vars(`Job Title`), ncol = 5) + 
#     ggtitle("Pie Chart of Degrees by Each Job Title") + 
#     theme(legend.position = "bottom")
```

```{r q3-d}
print_pie_chart <- function(job){
    edu <- sort(table(dat[dat$Q5 == job,]$Q4), decreasing = T) %>% 
        as.data.frame() %>%
        as.tibble()
    edu <- rename(edu, `Degree` = Var1)
    print(ggplot(edu, aes(x="", y=Freq, fill=`Degree`)) +
        geom_bar(stat="identity", width=1, color="white") +
        coord_polar("y", start=0) + 
        ggtitle(job) +
        theme_void())
}
jobs <- unique(dat$Q5)

print_pie_chart("Software Engineer")
print_pie_chart("Data Engineer")
```

## Gender Gap in Income

Is there a significant income gap between genders for these jobs?

### What about at an aggregate level? 
```{r q4}
# Q5

table(dat$Q2, dat$Q25)
```

As seen above at an aggregate level, for salary levels high and above there are a total of 799 men and only 143 women. The gender map is apparent. 

### By groups aggregated depending on years of experience? 

```{r q4-a}
for (agegroup in unique(dat$Q6)){
    if (agegroup != "I have never written code"){
        print(agegroup)
        print(table(dat$Q2[dat$Q6 == agegroup], dat$Q25[dat$Q6 == agegroup]))
    }
}
```

## Skillset vs Income

What is the typical skill set for these jobs? How does it affect the pay rate?

Here the key skill is defined as a skill that has been acquired more than 10% people under certain job title.

```{r q5-skill}
# Q5

skill.set <- job.dat %>% 
    filter(Q5 != "Other") %>%
    select(c(Q5, starts_with("Q7_"), starts_with("Q9_"), 
             starts_with("Q12_"), starts_with("Q14_"),
             starts_with("Q16_"), starts_with("Q17_"),
             starts_with("Q18_"), starts_with("Q19_"),
             salary_lb)) %>%
    mutate(Total = "`Total`") %>%
    gather("fake_key", "skillset", 
           -c(Q5, salary_lb), na.rm = T) %>%
    filter(!skillset %in% c("None", "Other")) %>%
    rename(title = Q5) %>%
    group_by(title, skillset) %>%
    summarise(n = n(), 
              salary_mean = round(mean(salary_lb, na.rm = T)),
              salary_sd = round(sd(salary_lb, na.rm = T)),
              .groups = "drop") %>%
    group_by(title) %>%
    mutate(prop = round(n * 100 / max(n), 1)) %>%
    filter(prop >= 0.1) %>%
    select(-n) %>%
    arrange(title, desc(prop))

datatable(skill.set, filter = 'top', width = 800)
```

From the table, huge salary variances make it very hard to tell whether a skill will increase the salary or not. 

## Correlation between Industry and Job

Is there a certain correlation between industry and the need for these jobs?

```{r q6-ind}
# Q6

industry.dat <- job.dat %>%
    filter(Q5 != "Student") %>%
    select(Q5, Q20, salary_lb, salary_ub) %>%
    mutate(Q20 = case_when(
        Q20 == "Academics/Education" ~ "Academics",
        Q20 == "Accounting/Finance" ~ "Finance", 
        Q20 == "Computers/Technology" ~ "Computers",
        Q20 == "Medical/Pharmaceutical" ~ "Medical",
        Q20 == "Online Service/Internet-based Services" ~ "Internet",
        TRUE ~ Q20
    )) %>%
    filter(Q20 %in% c("Academics", "Finance", "Computers",
                      "Medical", "Internet"))

p <- industry.dat %>% 
    count(Q5, Q20) %>%
    mutate(Q20 = fct_reorder(Q20, n, .fun="sum")) %>%
    rename(title=Q5, Industry=Q20, count=n) %>%
    ggplot(aes(x=title, y=count)) +
    geom_bar(stat = "identity") +
    coord_flip() +
    facet_wrap(~ Industry) +
    labs(
        title = "Users' work industry",
        caption = glue("Author: celeritasML
                   Source: Kaggle")) +
    theme(axis.ticks.x = element_blank(),
          axis.title = element_blank(),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank())
ggplotly(p)
```

A prop.table version.

```{r q6-ind-table}
ind.dis <- industry.dat %>% 
    filter(Q5 != "Statistician") %>%
    rename(Title = Q5, Industry = Q20)
ind.dis %>%
    group_by(Industry) %>%
    count(Title) %>%
    mutate(Proportion = round(prop.table(n), 3)) %>%
    select(-n) %>%
    pivot_wider(names_from = Industry, values_from = Proportion)
```

Some hypothesis that use prop.test to examine:  

```{r q6-tests}
aca.dis <- ind.dis %>% filter(Industry == "Academics")
com.dis <- ind.dis %>% filter(Industry == "Computers")
fin.dis <- ind.dis %>% filter(Industry == "Finance")
int.dis <- ind.dis %>% filter(Industry == "Internet")
med.dis <- ind.dis %>% filter(Industry == "Medical")

ds.dis <- ind.dis %>% filter(Title == "Data Scientist")
da.dis <- ind.dis %>% filter(Title == "Data Analyst")
de.dis <- ind.dis %>% filter(Title == "Data Engineer")

my.prop.test <- function(target.dis, other.dis, title) {
    target <- sum(target.dis$Title == title)
    other <- sum(other.dis$Title == title)
    t.res <- prop.test(c(target, other),
          c(nrow(target.dis), nrow(other.dis)),
          alternative = "greater",
          correct = TRUE)
    return(paste0("X-squared = ", round(t.res$statistic,1), 
                  ", p-value = ", formatC(t.res$p.value, 
                                          format = "e", digits = 2),
                  ", 95% CI = (", round(t.res$conf.int[1], 2), 
                  ", ", round(t.res$conf.int[2], 2),"), ",
                  if_else(t.res$p.value < 0.05, "H0 rejected", 
                          "can't reject H0")))
}

my.prop.test2 <- function(ind.dis, title1, title2) {
    target <- sum(ind.dis$Title == title1)
    other <- sum(ind.dis$Title == title2)
    t.res <- prop.test(c(target, other),
          c(nrow(ind.dis), nrow(ind.dis)),
          alternative = "greater",
          correct = TRUE)
    return(paste0("X-squared = ", round(t.res$statistic,1), 
                  ", p-value = ", formatC(t.res$p.value, 
                                          format = "e", digits = 2),
                  ", 95% CI = (", round(t.res$conf.int[1], 2), 
                  ", ", round(t.res$conf.int[2], 2),"), ",
                  if_else(t.res$p.value < 0.05, "H0 rejected", 
                          "can't reject H0")))
}
```

(1) $H_a$: `Data Analyst` is more needed in `Academics` Industry.

To `Computers`: `r my.prop.test(aca.dis, com.dis, "Data Analyst")`.  
To `Finance`: `r my.prop.test(aca.dis, fin.dis, "Data Analyst")`.  
To `Internet`: `r my.prop.test(aca.dis, int.dis, "Data Analyst")`.  
To `Medical`:  `r my.prop.test(aca.dis, med.dis, "Data Analyst")`.

**TRUE**, `Data Analyst` is more needed in `Academics` Industry.

(2) $H_a$: `Data Scientist` is more needed in `Medical` Industry.

To `Academics`: `r my.prop.test(med.dis, aca.dis, "Data Scientist")`.  
To `Computers`: `r my.prop.test(med.dis, com.dis, "Data Scientist")`.  
To `Finance`: `r my.prop.test(med.dis, fin.dis, "Data Scientist")`.  
To `Internet`: `r my.prop.test(med.dis, int.dis, "Data Scientist")`.  

**FALSE**, `Data Analyst` has similar demand in these industries.

(3) $H_a$: `Software Engineer` is more needed in `Computers` Industry.

To `Academics`: `r my.prop.test(com.dis, aca.dis, "Software Engineer")`.  
To `Finance`: `r my.prop.test(com.dis, fin.dis, "Software Engineer")`.  
To `Internet`: `r my.prop.test(com.dis, int.dis, "Software Engineer")`.  
To `Medical`:`r my.prop.test(com.dis, med.dis, "Software Engineer")`.  

**TRUE**, `Software Engineer` is more needed in `Computers` Industry.

(4) $H_a$: `Machine Learning Engineer` is more needed in `Internet` Industry.

To `Academics`: `r my.prop.test(com.dis, aca.dis, "Machine Learning Engineer")`.  
To `Computers`: `r my.prop.test(med.dis, com.dis, "Machine Learning Engineer")`.  
To `Finance`: `r my.prop.test(com.dis, fin.dis, "Machine Learning Engineer")`.  
To `Medical`: `r my.prop.test(com.dis, med.dis, "Machine Learning Engineer")`.  

**FALSE**, The demand of `Machine Learning Engineer` in `Internet` Industry is not significantly greater than in `Academics` and `Computers`.

(5) $H_a$: Overall, `Data Scientist` is more needed than other jobs.

To `Data Analyst`: `r my.prop.test2(ind.dis, "Data Scientist", "Data Analyst")`.  
To `Data Engineer`: `r my.prop.test2(ind.dis, "Data Scientist", "Data Engineer")`.  
To `Software Engineer`: `r my.prop.test2(ind.dis, "Data Scientist", "Software Engineer")`.  
To `ML Engineer`: `r my.prop.test2(ind.dis, "Data Scientist", "Machine Learning Engineer")`.  

**TRUE**, `Data Scientist` is more needed than other 4 jobs.

Salaries for these jobs vary widely from industry to industry.

```{r q6-ind-sal}
p <- industry.dat %>% 
    filter(salary_lb != 1000000) %>%
    mutate(Q20 = fct_reorder(Q20, salary_lb, .fun='length')) %>%
    ggplot(aes(x=Q20, y=salary_lb)) +
    geom_boxplot() +
    coord_flip() +
    facet_wrap(~ Q5) +
    labs(
        title = "Users' salary vs industry",
        caption = glue("Author: celeritasML
                   Source: Kaggle")) +
    theme(axis.ticks.x = element_blank(),
          axis.text.x = element_blank(),
          axis.title = element_blank(),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank())
ggplotly(p, tooltip="text")
```

## Languages and IDEs

What programming languages and IDEs do they use?

Survey questions Q7 (daily-used programming language), Q9 (IDE).

In the heatmap of programming language below, some trends can be found:

 - `Python`, `R`, and `SQL` are three most popular language in Kaggle community.
 - `Statisticians` use `R` more than `Python` and `SQL`.
 - `C++`, `Java`, `Javascript` are mainly used by `Software Engineer`.
 - `Julia` and `Swift` are barely used by Kaggle users.

```{r q7-lan}
# Q7

programming <- job.dat %>% 
    select(c(Q5, starts_with("Q7_"))) %>%
    gather("fake_key", "language", -Q5, na.rm = T) %>%
    rename(title = Q5) %>%
    select(-fake_key) %>%
    filter(!language %in% c("None", "Other")) %>%
    count(title, language, .drop = FALSE) %>% 
    complete(title, language) %>%
    replace_na(list(n = 0)) %>%
    group_by(title) %>%
    mutate(prop = prop.table(n))

p <- programming %>% 
    mutate(text = paste0("Language: ", language, "\n", 
                         "Job title: ", title, "\n", 
                         "Count: ", n, "\n",
                         "Proportion: ", round(prop, 3))) %>%
    ggplot(aes(language, title, fill=prop, text=text)) +
    geom_tile() +
    scale_fill_gradient(low="white", high="royalblue") +
    labs(
        title = "Users' favorite programming language",
        caption = glue("Author: celeritasML
                   Source: Kaggle")) +
    theme(axis.ticks.x = element_blank(),
          axis.text.x = element_text(angle=90, hjust=1),
          axis.title = element_blank(),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank())
ggplotly(p, tooltip="text")
```

In the heatmap of IDE below, some trends can be found:

 - `Jupyter Notebook` is the most popular IDE for Kaggle users.
 - `MATLAB` is barely used in the industry(last 5 jobs).
 - `Statisticians` use `RStudio` the most.
 - `VSCode` is overall the second common-used IDE.

```{r q7-ide}
ide <- job.dat %>% 
    select(c(Q5, starts_with("Q9_"))) %>%
    gather("fake_key", "IDE", -Q5, na.rm = T) %>%
    rename(title = Q5) %>%
    select(-fake_key) %>%
    mutate(IDE = case_when(
        IDE == "Visual Studio Code (VSCode)" ~ "VSCode",
        IDE == "Jupyter (JupyterLab, Jupyter Notebooks, etc)" ~ "Jupyter Notebook",
        TRUE ~ IDE
    )) %>%
    filter(!IDE %in% c("None", "Other")) %>%
    count(title, IDE, .drop = FALSE) %>% 
    complete(title, IDE) %>%
    replace_na(list(n = 0)) %>%
    group_by(title) %>%
    mutate(prop = prop.table(n))

p <- ide %>% 
    mutate(text = paste0("IDE: ", IDE, "\n", 
                         "Job title: ", title, "\n", 
                         "Count: ", n, "\n",
                         "Proportion: ", round(prop, 3))) %>%
    ggplot(aes(IDE, title, fill=prop, text=text)) +
    geom_tile() +
    scale_fill_gradient(low="white", high="royalblue") +
    labs(
        title = "Users' favorite IDE",
        caption = glue("Author: celeritasML
                   Source: Kaggle")) +
    theme(axis.ticks.x = element_blank(),
          axis.text.x = element_text(angle=90, hjust=1),
          axis.title = element_blank(),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank())
ggplotly(p, tooltip="text")
```

From two plots above, while Python becomes the most popular programming language for Kaggle users, Jupyter Notebook (a Python IDE) also becomes the most commonly used IDE. Are there any correlation between them?

```{r q7-py-jupyter}
pj.dat <- job.dat %>% 
        select(c(Q5, Q7_Part_1, Q9_Part_1, Q9_Part_11)) %>% 
        rename(Title = Q5) %>%
        mutate(Python = !is.na(Q7_Part_1),
               Jupyter = !is.na(Q9_Part_1) | !is.na(Q9_Part_11)) %>%
        select(Title, Python, Jupyter) %>%
        pivot_longer(cols = c(Python, Jupyter)) %>%
        filter(value == TRUE)
py.ju.test <- function(excludes) {
    pj.t.dat <- pj.dat %>% filter(!Title %in% excludes)
    print(table(pj.t.dat$name, pj.t.dat$Title))
    pj.t.res <- chisq.test(pj.t.dat$name, pj.t.dat$Title, 
                           simulate.p.value = T)
    return(paste0("X-squared = ", round(pj.t.res$statistic, 1), 
                  ", p-value = ", round(pj.t.res$p.value, 2),
                  if_else(pj.t.res$p.value < 0.05, ", H0 rejected", 
                          ", can't reject H0")))
}
table(pj.dat$Title, pj.dat$name)
```

$H_a$: The distribution of Python is significantly different from the distribution of Jupyter Notebook.

All job titles: `r py.ju.test(c())`.  
Without `Statistician`: `r py.ju.test(c("Statistician"))`.  
Without `Statistician`, `Student`: `r py.ju.test(c("Statistician", "Student"))`.  

**FALSE**. There is no significant difference between the distribution of Python and the distribution of Jupyter Notebook among jobs. In the other words, most people who use Python also use Jupyter Notebook as the IDE.

## Learning Sources

Where do they get and share the knowledge?

Survey questions Q39 (share and deploy), Q40 (learning resources), Q42 (Media sources).

In the heatmap of learning platform below, some trends can be found:

 - `Coursera` is the most popular learning platform outside the University.
 - `Statisticans` learn knowledge mainly from `University` course.

```{r q8-learn}
# Q8

learning_platform <- job.dat %>%
    select(c(Q5, starts_with("Q40_"))) %>%
    gather("fake_key", "learning", -Q5, na.rm = T) %>%
    rename(title = Q5) %>%
    select(-fake_key) %>%
    mutate(learning = case_when(
        learning == "Cloud-certification programs (direct from AWS, Azure, GCP, or similar)" ~ "Cloud-certif Programs",
        learning == "University Courses (resulting in a university degree)" ~ "University",
        TRUE ~ learning
    )) %>%
    filter(!learning %in% c("None", "Other")) %>%
    count(title, learning, .drop = FALSE) %>%
    complete(title, learning) %>%
    replace_na(list(n = 0)) %>%
    group_by(title) %>%
    mutate(prop = prop.table(n))

p <- learning_platform %>%
    mutate(text = paste0("Platform: ", learning, "\n",
                         "Job title: ", title, "\n",
                         "Count: ", n, "\n",
                         "Proportion: ", round(prop, 3))) %>%
    ggplot(aes(learning, title, fill=prop, text=text)) +
    geom_tile() +
    scale_fill_gradient(low="white", high="royalblue") +
    labs(
        title = "Users' favorite learning platforms",
        caption = glue("Author: celeritasML
                   Source: Kaggle")) +
    theme(axis.ticks.x = element_blank(),
          axis.text.x = element_text(angle=90, hjust=1),
          axis.title = element_blank(),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank())
ggplotly(p, tooltip="text")
```

In the heatmap of sharing platform below, some trends can be found:

 - `GitHub` is the most popular sharing platform.
 - `Kaggle`, `Colab`, `Personal Blog` are used by a small proportion of Kaggle Users.
 - Other platforms listed on the questionare are barely used by Kaggle Users.
 - A large proportion of people don't share their works to public.
 - None of `Student` answered this question.

```{r q8-share}
share_deploy <- job.dat %>% 
    select(c(Q5, starts_with("Q39_"))) %>%
    gather("fake_key", "share", -Q5, na.rm = T) %>%
    rename(title = Q5) %>%
    select(-fake_key) %>%
    mutate(share = case_when(
        share == "I do not share my work publicly" ~ "\'PRIVATE\'",
        TRUE ~ share
    )) %>%
    filter(!share %in% c("Other")) %>%
    count(title, share, .drop = FALSE) %>% 
    complete(title, share) %>%
    replace_na(list(n = 0)) %>%
    group_by(title) %>%
    mutate(prop = prop.table(n))

p <- share_deploy %>% 
    mutate(text = paste0("Platform: ", share, "\n", 
                         "Job title: ", title, "\n",
                         "Count: ", n, "\n",
                         "Proportion: ", round(prop, 3))) %>%
    ggplot(aes(share, title, fill=prop, text=text)) +
    geom_tile() +
    scale_fill_gradient(low="white", high="royalblue") +
    labs(
        title = "Users' favorite share platforms",
        x = "",
        y = "",
        caption = glue("Author: celeritasML
                   Source: Kaggle")) +
    theme(axis.ticks.x = element_blank(),
          axis.text.x = element_text(angle=90, hjust=1),
          axis.title = element_blank(),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank())
ggplotly(p, tooltip="text")
```

In the heatmap of media source below, some trends can be found:

 - `Blog`, `Kaggle`, `YouTube` are three most popular media source.
 - Many `Statisticians` use `Journal Publication` as their media sources.

```{r q8-media}
media_source <- job.dat %>% 
    select(c(Q5, starts_with("Q42_"))) %>%
    gather("fake_key", "media", -Q5, na.rm = T) %>%
    rename(title = Q5) %>%
    select(-fake_key) %>%
    filter(!media %in% c("None", "Other")) %>%
    count(title, media, .drop = FALSE) %>% 
    complete(title, media) %>%
    replace_na(list(n = 0)) %>%
    group_by(title) %>%
    mutate(prop = prop.table(n)) %>%
    separate(media, into = c("media", "media_suffix"), sep = " \\(")

p <- media_source %>% 
    mutate(text = paste0("Platform: ", media, "\n", 
                         "Job title: ", title, "\n", 
                         "Count: ", n, "\n", 
                         "Proportion: ", round(prop, 3))) %>%
    ggplot(aes(media, title, fill=prop, text=text)) +
    geom_tile() +
    scale_fill_gradient(low="white", high="royalblue") +
    labs(
        title = "Users' favorite media source",
        caption = glue("Author: celeritasML
                   Source: Kaggle")) +
    theme(axis.ticks.x = element_blank(),
          axis.text.x = element_text(angle=90, hjust=1),
          axis.title = element_blank(),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank())
ggplotly(p, tooltip="text")
```

Special Job Title: `Statistician`

From all analysis above, `Statistician` shows many different preference than other jobs. For example:

 - Only a few people identify themselves as `Statistician`.
 - They mostly work in `Academic` industry.
 - They use `R` more frequently, also they use `RStudio` a lot.
 - They learn most knowledge from `University Courses`.
 - Many of them use `Journal Publication` to get report on data science topics.

Based on these features, `Statisticians` are very likely to be a job title specially for professors and researchers in academic institutions (i.e. universities). 

## Popular Packages

As for the popular packages, one can easily give some examples that no strangers to data science practitioners. But the real problem is, what can be learned from this? Some conditional probabilities will highlight the findings.

The data is a subset of survey answer to Q14 and Q16, which emphasize on visualization libraries and machine learning related libraries respectively. A set of niche job titles are excluded due to limited number of responses, such as "DBA/Database Engineer", "Currently not employed" and "Statistician." The remaining valid responses are visualized as barplots by group to show the count of responses that are familiar with those packages.

Additionally, as the number of machine learning libraries covered in the survey exceeds a reasonable number of meaning palette colors, only the top 10 libraries are kept in the visualization.

```{r q9-viz}
# Q9

viz_lib <- dat %>%
  select(Q5, starts_with("Q14")) %>%
  select(-c(Q14_Part_11, Q14_OTHER))

viz_lib <- viz_lib %>%
  pivot_longer(cols=starts_with("Q14")) %>%
  select(-name) %>%
  drop_na() %>%
  filter(!(Q5 %in% c("Other", "DBA/Database Engineer",
                   "Developer Relations/Advocacy",
                   "Currently not employed",
                   "Statistician", "Product Manager")))

ggplot(viz_lib) +
  geom_bar(aes(y = value, fill = value)) +
  facet_wrap(~ Q5) +
  scale_fill_brewer(palette = "Spectral") +
  labs(
    title = "Data science practitioners' favorite viz libraries",
    x = "",
    y = "",
    caption = glue("Author: celeritasML
                   Source: Kaggle")
  ) +
  theme(
        axis.ticks.x = element_line(),
        axis.ticks.y = element_blank(),
        axis.text.x = element_text(size = 6),
        axis.text.y = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank())
```
```{r q9-ml}
ml_lib <- dat %>%
  select(Q5, starts_with("Q16")) %>%
  select(-c(Q16_Part_17, Q16_OTHER))

ml_lib <- ml_lib %>%
  pivot_longer(cols=starts_with("Q16")) %>%
  select(-name) %>%
  drop_na() %>%
  filter(!(Q5 %in% c("Other", "DBA/Database Engineer",
                   "Developer Relations/Advocacy",
                   "Currently not employed",
                   "Statistician", "Product Manager")))

top_10_ml <- ml_lib %>%
  group_by(value) %>%
  summarize(count = n()) %>%
  arrange(desc(count)) %>%
  top_n(10)

ml_lib <- ml_lib %>%
  filter(value %in% top_10_ml$value)

ggplot(ml_lib) +
  geom_bar(aes(y = value, fill = value)) +
  facet_wrap(~ Q5) +
  scale_fill_brewer(palette = "Spectral") +
  labs(
    title = "Data science practitioners' favorite ML libraries",
    x = "",
    y = "",
    caption = glue("Author: celeritasML
                   Source: Kaggle")
  ) +
  theme(
        axis.ticks.x = element_line(),
        axis.ticks.y = element_blank(),
        axis.text.x = element_text(size = 6),
        axis.text.y = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank())
```

There are some obvious results worth mentioning. But before that, it is important to reiterate that the comparison is based on the calculation of some conditional probabilities according to Bayes' Theorem.

$$P(A|B) = \frac{P(B|A)\cdot P(A)}{P(B)} = \frac{P(A \cap  B)}{P(B)}$$

where $A$ and $B$ are events.

```{r}
viz_lib %>%
  group_by(Q5) %>%
  table() %>%
  addmargins() %>%
  print()

ml_lib %>%
  group_by(Q5) %>%
  table() %>%
  addmargins() %>%
  print()
```

- Both **Data Scientists** and **Students** are the two major respondents to package preferences/familiarity questions.

$$
\begin{split}
  P(\text{Student}|\text{Visualization}) &= \frac{739}{3923} \approx 0.1884\\
  P(\text{Data Scientist}|\text{Visualization}) &= \frac{1165}{3923} \approx 0.2970 \\
  P(\text{Student}|\text{Machine Learning}) &=\frac{732}{4038} \approx 0.1813\\
  P(\text{Data Scientist}|\text{Machine Learning}) &=\frac{1364}{4038} \approx 0.3378
\end{split}
$$

- Python-related packages or packages with Python support are overall more popular among the responses.
  - For example, `Altair`, `bokeh`, `geoplotlib`, `leaflet/folium`, `matplotlib`, `plotly/plotly express`, `seaborn` are categorized as Python-wise, while `D3.js`, `ggplot2`, `leaflet/folium`, `plotly/plotly express` and `shiny` are categorized as R-wise. Note that there are some overlapping among these packages since they do have multiple language support.
  - Similarly, `Huggingface`, `Keras`, `LightGBM`, `Prophet`, `PyTorch`, `scikit-learn`, `TensorFlow`, `Xgboost` are Python compatible. `Caret`, `Keras`, `LightGBM`, `Prophet`, `TensorFlow`, `tidymodels` and `Xgboost` and usable in R.

$$
\begin{split}
P(\text{Python}|\text{Visualization}) &= (33+115+96+80+1247+525+861)/3923 \approx 0.7538\\ 
P(\text{R}|\text{Visualization}) &= (132+622+80+525+212)/3923 \approx 0.4005\\
P(\text{Python}|\text{Machine Learning}) &= (132+549+187+93+527+1125+666+506)/4038 \approx 0.9373\\ 
P(\text{R}|\text{Machine Learning}) &= (148+549+187+93+666+105+506)/4038 \approx 0.5582\\
\end{split}
$$

There are more to be discovered depending on the perspectives.

## The Loyalty to AWS

As the demand for computational power increases along with the amount of data involved in data science industry, cloud computing is a topic that any data science practitioners cannot avoid. The market is majorly occupied by **Amazon, Google** and **Microsoft**.

Specifically, will a user's preference for a cloud computing platforms affect his or her preferences for other tools? For example, an AWS EC2 dedicated user will actually prefer AWS S3 over other products.

A list of AWS services are selected for this hypothesis, including EC2, S3, EFS, SageMaker, RedShift, Aurora, RDS and DynamoDB. The data covers the following survey questions:

- Survey question Q29-A: computing products (Part_1)
- Survey question Q30: Storage (Part_3, Part_4)
- Survey question Q31-A: ML products (Part_1)

The null hypothesis to be tested is:

- $H_0:$ The proportion of people who use EC2 is independent of the people who uses S3 (this could be replaced by any other AWS services.)
- $H_a:$ The proportion of people who use EC2 is dependent on the people who use S3.

The hypothesis is tested by checking the $\chi^2$ as the statistic.

$$\chi^2=\sum\frac{(O_i-E_i)^2}{E_i}$$
The following three tests are conducted between:

- EC2 vs S3
- EC2 vs EFS
- EC2 vs SageMaker

```{r q10}
# Q10

aws_user <- tibble(
  ec2 = dat$Q29_A_Part_1,
  s3 = dat$Q30_A_Part_3,
  efs = dat$Q30_A_Part_4,
  sagemaker = dat$Q31_A_Part_1,
  redshift = dat$Q32_A_Part_11,
  aurora = dat$Q32_A_Part_12,
  rds = dat$Q32_A_Part_13,
  dynamodb = dat$Q32_A_Part_14
  
) %>%
  mutate(ec2 = if_else(is.na(ec2), 0, 1),
         s3 = if_else(is.na(s3), 0, 1),
         efs = if_else(is.na(efs), 0, 1),
         sagemaker = if_else(is.na(sagemaker), 0, 1),
         redshift = if_else(is.na(redshift), 0, 1),
         aurora = if_else(is.na(aurora), 0, 1),
         rds = if_else(is.na(rds), 0, 1),
         dynamodb = if_else(is.na(dynamodb), 0, 1))
```

```{r q10-a}
chisq.test(aws_user$ec2, aws_user$s3)
```

```{r q10-b}
chisq.test(aws_user$ec2, aws_user$efs)
```

```{r q10-c}
chisq.test(aws_user$ec2, aws_user$sagemaker)
```

Since all three p-values are significantly smaller than the common critical value 0.05, that means the null hypothesis is rejected. That means the user group of EC2 and all three AWS services' user groups (S3, EFS and SageMaker) are not dependent.

A possible explanation is that EC2 is the core service of AWS, and most of other AWS services rely on running an elastic computing instance. That implies a data science practitioner who frequently uses EC2 will be more likely to use other AWS services.

## Cloud Computing Platforms Preferences

Another related question addressed is to compare the market occupancy rate of the top 3 cloud computing providers. But in the survey, the market occupancy rate is instantiated as the using rate among survey takers.

Again, two hypotheses are given:

- $H_0: p_A=p_B$
- $H_a:p_A\not=p_B$

where $A$ and $B$ can be replaced by AWS, Azure, or GCP, and $n_A$, $n_B$ are sample size of group $A$ and $B$ respectively.

The test statistic (z-statistic) can be calculated as follow:

$$z=\frac{p_A-p_B}{\sqrt{p(1-p)/n_A+p(1-p)/n_B}}$$

```{r q11}
# Q11

cloud_comp <- tibble(
  aws_usage = dat$Q27_A_Part_1,
  azure_usage = dat$Q27_A_Part_2,
  gcp_usage = dat$Q27_A_Part_3
) %>%
  mutate(aws_usage = if_else(is.na(aws_usage), FALSE, TRUE),
         azure_usage = if_else(is.na(azure_usage), FALSE, TRUE),
         gcp_usage = if_else(is.na(gcp_usage), FALSE, TRUE))
```

```{r q11-a}
prop.test(c(sum(cloud_comp$aws_usage), sum(cloud_comp$azure_usage)),
          c(nrow(cloud_comp), nrow(cloud_comp)),
          alternative = "two.sided",
          correct = TRUE)
```

```{r q11-b}
prop.test(c(sum(cloud_comp$azure_usage), sum(cloud_comp$gcp_usage)),
          c(nrow(cloud_comp), nrow(cloud_comp)),
          alternative = "two.sided",
          correct = TRUE)
```

```{r q11-c}
prop.test(c(sum(cloud_comp$gcp_usage), sum(cloud_comp$aws_usage)),
          c(nrow(cloud_comp), nrow(cloud_comp)),
          alternative = "two.sided",
          correct = TRUE)
```

The results show that only the p-value of second test is greater than the common critical value 0.05. That means:

- Reject $H_0$ that the using rates of AWS and Microsoft Azure are the same.
- Fail to reject $H_0$ that the using rates of Microsoft Azure and Google Cloud Platform are the same.
- Reject $H_0$ that the using rates of Google Cloud Platform and AWS are the same.

Therefore, to simplify the result, AWS is leading the market (a lot), while Microsoft Azure and Google Cloud Platform are pretty much the same in terms of using rate.

Another perspective of interpretation is to look at the 95% confidence interval of the hypothesis test. For instance, the test between AWS usage and Azure usage gives a confidence interval of $(0.076, 0.119)$. So it's 95% confident that the true difference is between such an interval.

## Linear Regression on Salary

Honestly speaking, the best way to approach this question is to apply a logistic regression on the the dependent variable `salary` due to its categorical nature. However, in order to experiment a linear model, we take **a bold move to transform the `salary` back to continuous by randomly sampling values in between the pay levels**. Three typical languages used by data science practitioners are included as well: Python, R and SQL.

A quick glimpse at the response will uncover the fact that the `salary` is not normally distributed. In fact, it is right-skewed with a long tail. A log-transformation on `salary` will not fully fix this issue, but indeed make the linear model fitting easier.

```{r q12-lm}
# Q12

lm_dat <- dat %>%
  select(Q1, Q2, Q4, Q5, Q6,
         Q7_Part_1, Q7_Part_2, Q7_Part_3,
         Q25) %>%
  drop_na(Q25) %>%
  rename(age_group = Q1,
         gender = Q2,
         education = Q4,
         job = Q5,
         experience = Q6,
         salary = Q25,
         python = Q7_Part_1,
         r = Q7_Part_2,
         sql = Q7_Part_3)

set.seed(511)

# - poverty: below the federal minimum wage
# - low: 40,000 to 79,999
# - medium: 80,000 to 124,999
# - high: 125,000 to 199,999
# - very high: 200,000 to 499,999
# - highest: >= 500,000

lm_dat <- lm_dat %>%
  rowwise() %>%
  mutate(salary = case_when(
    salary == "poverty" ~ sample(1:39999, 1),
    salary == "low" ~ sample(40000:79999, 1),
    salary == "medium" ~ sample(80000:124999, 1),
    salary == "high" ~ sample(125000:199999, 1),
    salary == "very high" ~ sample(200000:499999, 1),
    salary == "highest" ~ sample(500000:2000000, 1)
  )) %>%
  ungroup()

ggplot(lm_dat, aes(x=salary)) +
  geom_histogram(binwidth = 5000) +
  labs(
        title = "Histogram of salaries",
        subtitle = "raw data without transformation",
        caption = glue("Author: celeritasML
                   Source: Kaggle"))

ggplot(lm_dat, aes(x=log(salary))) +
  geom_histogram(binwidth = 0.05) +
  labs(
        title = "Histogram of log(salary)",
        subtitle = "after log-transformation on response",
        caption = glue("Author: celeritasML
                   Source: Kaggle"))

lm_dat <- lm_dat %>%
  mutate(python = if_else(is.na(python), 0, 1),
         r = if_else(is.na(r), 0, 1),
         sql = if_else(is.na(sql), 0, 1))
```

Afterwards, the function `MASS:stepAIC()` is used to do a stepwise regression in order to find the suitable predictive variables. Note that the selection criterion is AIC, and the model starts with the maximal model with all the predictors and some manually added interactions, including:

- Interactions between languages, `sql:r`, `sql:python`, `r:python`, `python:r:sql`.
- Interactions between jobs and languages, `job:sql`, `job:r`, `job:python`.
- Interactions between jobs and other variables, `job:experience`, `job:education`.

```{r q12-lm-model}
model1 <- lm(log(salary) ~ . + sql:r + sql:python + r:python + python:r:sql +
               job:sql + job:r + job:python + job:experience +
               job:education, data=lm_dat)
model2 <- MASS::stepAIC(model1, trace = FALSE)
```

Finally, the automatic procedure of model fitting returns the following model as the result:

$$\log(\text{salary})=\beta_0+\beta_1 \text{age_group} +\beta_2\text{gender}+
\beta_3\text{job}+\beta_4\text{experience}\\ +\beta_5\text{python}+\beta_6\text{r}+\beta_7\text{sql}\\
+\beta_8\text{python}\cdot\text{sql} + \beta_9\text{python}\cdot\text{r} +
\beta_{10}\text{job}\cdot\text{python} + \epsilon
$$

```{r q12-final-model-anova}
anova(model2)
```

```{r q12-final-model-summary}
summary(model2)
```

By looking at both the `summary()` table and `anova()` table, the following can be concluded as the takeaways:

- The predictive variables `age_group`, `gender`, `job` and `experience` are significant while the three languages `python`, `r` and `sql` are not significant alone.
- However, the interaction of `python:sql`, `python:r` and `job:python` are playing some roles in predicting the `log(salary)` as the response.
- Generally speaking, the `age_group` shows new graduate (22-24) and people past 70+ tend to have lower salaries.
- Comparing to the baseline `gender`, which is `Male`, other gender options lead to lower salaries.
- The `experience` years and `salary` are positively related.
- Some job titles with `python` as a skillset can significantly relate to `salary` changes. Although it is still confusing to see the combination of `Developer Relations/Advocacy:python` has quite a lot negative effect on `salary`.

In general, this linear model is built upon lots of compromising assumptions, so that the overall power to explain the deviations in not ideal. Nevertheless, it still shows some insights about the data.


# Conclusions

- findings as bullet points:

1. We don't have a one-sentence conclusion to summarize all the findings above. However, what we learned from this project is more about...
2. gender inequality
3. ...

- Besides the loose collection of findings above, we believe that (ML/cloud computing) is the general trend of data science in the next couple of years.
- More to discover from the survey data. Comparison with historical data blah blah blah.
- Limitation. (from our proposal, last paragraph). Kaggle users != ds practitioners; lots of students -> not representative enough to reflect the reality in the industry.


# Appendix: R scripts

```{r ref.label=knitr::all_labels(), echo=TRUE, eval=FALSE}
```

See footnote 1[^1].

[^1]: See GitHub repository [tufte](https://github.com/rstudio/tufte).
