---
title: "report_q12345"
author: "Ercong Luo"
date: "11/29/2021"
output: html_document
---

Questionnaire Answer Choices: 
https://github.com/CeleritasML/kaggle-ml-survey/blob/main/data/supplementary_data/kaggle_survey_2021_answer_choices.pdf

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# invalidate cache when the package version changes
knitr::opts_chunk$set(tidy = FALSE, cache.extra = packageVersion('tufte'))
options(htmltools.dir.version = FALSE)

if (!require("pacman")) install.packages("pacman")
pacman::p_load(tidyverse, ggthemes, latex2exp, glue,
               ggmosaic)

set.seed(511)
```

```{r data cleaning}

col_names <- names(read_csv(
  "data/kaggle_survey_2021_responses.csv",
  n_max=0))
dat <- read_csv(
  "data/kaggle_survey_2021_responses.csv",
  col_names = col_names, skip=2)

dat <- dat %>%
  filter(Q3=="United States of America" )
```
## Q1: Job Titles Distributions

What percentage of the survey respondents are working under these job titles? 

```{r}
jtitle <- sort(table(dat$Q5), decreasing = T) %>% 
    as.data.frame() %>%
    as.tibble()
jtitle <- rename(jtitle, `Job Title` = Var1)

ggplot(jtitle, aes(x="", y=Freq, fill=`Job Title`)) +
    geom_bar(stat="identity", width=1, color="white") +
    coord_polar("y", start=0) + 
    theme_void() 
```

## Q2: Salaries

What are the statistics on salaries for these job titles?

The US minimum wage is 7.25 per hour, multiply that for a total $260$ typical number of workdays a year:

$$7.25 \frac{\$}{\text{hour}} \cdot 24 \frac{\text{hours}}{\text{day}} \cdot 260\frac{\text{days}}{\text{year}} = 45240.$$

The new categories* for salaries will be:
- poverty: below the federal minimum wage
- low: 40,000 to 79,999
- medium: 80,000 to 124,999
- high: 125,000 to 199,999
- very high: 200,000 to 499,999
- highest: >= 500,000

*loosely based on US federal income tax brackets. 


```{r}
# To change salary categories into FACTOR dtype with descending labels
poverty   = c("$0-999", "1,000-1,999" , "2,000-2,999", "3,000-3,999", "4,000-4,999", "5,000-7,499", "7,500-9,999",
             "10,000-14,999", "15,000-19,999", "20,000-24,999", "25,000-29,999", "30,000-39,999") 
low       = c("40,000-49,999", "50,000-59,999", "60,000-69,999", "70,000-79,999")
medium    = c("80,000-89,999", "90,000-99,999", "100,000-124,999")
high      = c("125,000-149,999", "150,000-199,999")
very_high = c("200,000-249,999", "250,000-299,999", "300,000-499,999")
highest   = c("$500,000-999,999", ">$1,000,000")

dat$Q25[dat$Q25 %in% poverty] <- "poverty"
dat$Q25[dat$Q25 %in% low] <- "low"
dat$Q25[dat$Q25 %in% medium] <- "medium"
dat$Q25[dat$Q25 %in% high] <- "high"
dat$Q25[dat$Q25 %in% very_high] <- "very high"
dat$Q25[dat$Q25 %in% highest] <- "highest"
dat$Q25 = factor(dat$Q25, levels = c("poverty", "low", "medium", "high", "very high", "highest"), ordered = T)

data_side <- c("Data Scientist", "Data Analyst", "Business Analyst", "Data Engineer", "Statistician", "DBA/Database Engineer")
swe_side <- c("Software Engineer", "Machine Learning Engineer", "Program/Project Manager", "Product Manager") 
academia <- c("Student", "Other", "Research Scientist") 

dat[dat$Q5 %in% data_side & !is.na(dat$Q5) & !is.na(dat$Q25), ] %>%
  ggplot( aes(x=Q5, y=Q25, color = Q5)) +
    geom_count() + 
    ggtitle("Two-Way Salary Visualizations: Data-Oriented Jobs") +
    xlab("") +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))


dat[dat$Q5 %in% swe_side & !is.na(dat$Q5) & !is.na(dat$Q25), ] %>%
  ggplot( aes(x=Q5, y=Q25, color = Q5)) +
    geom_count() + 
    ggtitle("Two-Way Salary Visualizations: Engineering-Oriented Jobs") +
    xlab("") + 
    theme(axis.text.x = element_text(angle = 45, hjust = 1))

dat[dat$Q5 %in% academia & !is.na(dat$Q5) & !is.na(dat$Q25), ] %>%
  ggplot( aes(x=Q5, y=Q25, color = Q5)) +
    geom_count() + 
    ggtitle("Two-Way Salary Visualizations: Academic Jobs and Others") +
    xlab("") + 
    theme(axis.text.x = element_text(angle = 45, hjust = 1))

```


## Q3: Requried Education Levels

What levels of education are required for these job titles?

### Aggregate Level

```{r}
edu <- sort(table(dat$Q4), decreasing = T) %>% 
    as.data.frame() %>%
    as.tibble()
edu <- rename(edu, `Degree` = Var1)

ggplot(edu, aes(x="", y=Freq, fill=`Degree`)) +
    geom_bar(stat="identity", width=1, color="white") +
    coord_polar("y", start=0) + 
    theme_void() 
```


### Break down by Industry

## Q4: Gender Gap in Income

Is there a significant income gap between genders for these jobs?

### What about at an aggregate level? 
```{r}

```


### By groups aggregated depending on years of experience? 

### By industry? 

### By Geographical Location? 

## Q5: Skillset vs Income

### Cloud Computing Chops

### Programming Languages vs salary


What is the typical skill set for these jobs? How does it affect the pay rate?
