---
title: "Data Science Career Paths & Skillsets in 2021"
subtitle: "A glimpse of Kaggle’s State of Data Science and Machine Learning Survey"
author:
  - Jingsong Gao, jg2109@georgetown.edu
  - Ercong Luo, el890@georgetown.edu
  - Rui Qiu, rq47@georgetown.edu
date: "`r Sys.Date()`"
output:
  tufte::tufte_html:
    # tufte_variant: "envisioned"
    highlight: kate
---

```{r setup, include=FALSE}
# invalidate cache when the package version changes
knitr::opts_chunk$set(tidy = FALSE, cache.extra = packageVersion('tufte'))
options(htmltools.dir.version = FALSE)

if (!require("pacman")) install.packages("pacman")
pacman::p_load(tidyverse, ggthemes, latex2exp, glue,
               ggmosaic, plotly, stringr, DT)

set.seed(511)

ggplot2::theme_set(
  theme_fivethirtyeight() +
    theme(
        text = element_text(family = "Roboto Condensed"),
        title = element_text(size = 14),
        plot.subtitle = element_text(size = 12),
        plot.caption = element_text(size = 10),
        axis.title = element_text(size = 14),
        axis.text = element_text(size = 12),
        panel.grid.minor.x = element_blank()
    )
)
```

# Section 1: Introduction

Address the data science problem

# Section 2: Data Cleaning

```{r pkg, message=FALSE}
col_names <- names(read_csv(
  "data/kaggle_survey_2021_responses.csv",
  n_max=0))
dat <- read_csv(
  "data/kaggle_survey_2021_responses.csv",
  col_names = col_names, skip=2)

dat <- dat %>%
  filter(Q3=="United States of America" )

job.dat <- dat %>% 
    filter(Q5 %in% c("Data Analyst",
                     "Data Engineer",
                     "Data Scientist",
                     "Machine Learning Engineer",
                     "Software Engineer",
                     "Statistician",
                     "Student"))
```

# Section 3: The Analysis and Results

- List of questions.

```{marginfigure}
Since the one large data science problem can be divided into several smaller bits, we decide to combine the analysis and the interpretation sections together. And the questions will be addressed and discussed one by one.
```

```{marginfigure}
The following questions are about data science career paths.
```

## Q1: Job Titles Distributions

What percentage of the survey respondents are working under these job titles? 

## Q2: Salaries

What are the statistics on salaries for these job titles?

## Q3: Requried Education Levels

What levels of education are required for these job titles?

## Q4: Gender Gap in Income

Is there a significant income gap between genders for these jobs?

## Q5: Skillset vs Income

What is the typical skill set for these jobs? How does it affect the pay rate?

```{r q5-skill}
skill.set <- job.dat %>% 
    select(c(Q5, starts_with("Q7_"), starts_with("Q9_"), 
             starts_with("Q12_"), starts_with("Q14_"),
             starts_with("Q16_"), starts_with("Q17_"),
             starts_with("Q18_"), starts_with("Q19_"))) %>%
    mutate(Total = "TotalHelper") %>%
    gather("fake_key", "skillset", -Q5, na.rm = T) %>%
    filter(!skillset %in% c("None", "Other")) %>%
    rename(title = Q5) %>%
    count(title, skillset) %>%
    group_by(title) %>%
    mutate(prop = round(n / max(n), 3)) %>%
    filter(prop >= 0.1 & skillset != "TotalHelper") %>%
    select(-n) %>%
    arrange(title, desc(prop))

datatable(skill.set, filter = 'top')
```


## Q6: Correlation between Industry and Job

Is there a certain correlation between industry and the need for these jobs?

```{r q6-ind}
industry.dat <- job.dat %>%
    filter(Q5 != "Student") %>%
    select(Q5, Q20, Q25) %>%
    filter(Q20 %in% c("Academics/Education", 
                      "Accounting/Finance", 
                      "Computers/Technology",
                      "Insurance/Risk Assessment",
                      "Medical/Pharmaceutical",
                      "Online Service/Internet-based Services")) %>%
    mutate(Q25 = str_remove_all(Q25, "[$,]")) %>%
    mutate(Q25 = str_replace(Q25, ">1000000", "1000000-2000000")) %>%
    separate(Q25, into = c("salary_lb", "salary_ub"), sep = "-") %>%
    mutate(salary_lb = as.numeric(salary_lb)) %>%
    mutate(salary_ub = as.numeric(salary_ub))

p <- industry.dat %>% 
    count(Q5, Q20) %>%
    mutate(Q20 = fct_reorder(Q20, n, .fun="sum")) %>%
    rename(title=Q5, Industry=Q20, count=n) %>%
    ggplot(aes(x=Industry, y=count)) +
    geom_bar(stat = "identity") +
    coord_flip() +
    facet_wrap(~ title) +
    labs(
        title = "Users' work industry",
        caption = glue("Author: celeritasML
                   Source: Kaggle")) +
    theme(axis.ticks.x = element_blank(),
          axis.text.x = element_text(angle=90, hjust=1),
          axis.title = element_blank(),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank())
ggplotly(p)
```

```{r}
chisq.test(table(industry.dat$Q5, industry.dat$Q20))
```


```{r q6-ind-sal}
industry.dat %>% 
    mutate(Q20 = fct_reorder(Q20, salary_lb, .fun='length')) %>%
    ggplot(aes(x=Q20, y=salary_lb)) +
    geom_boxplot() +
    coord_flip() +
    facet_wrap(~ Q5) +
    labs(
        title = "Users' salary vs industry",
        caption = glue("Author: celeritasML
                   Source: Kaggle")) +
    theme(axis.ticks.x = element_blank(),
          axis.text.x = element_text(angle=90, hjust=1),
          axis.title = element_blank(),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank())
```


```{marginfigure}
The following questions are majorly about data science skills, tools and technologies.
```

## Q7: Languages and IDEs

What programming languages and IDEs do they use?

Survey questions Q7 (daily-used programming language), Q9 (IDE).

```{r q7-lan}
programming <- job.dat %>% 
    select(c(Q5, starts_with("Q7_"))) %>%
    gather("fake_key", "language", -Q5, na.rm = T) %>%
    rename(title = Q5) %>%
    select(-fake_key) %>%
    filter(!language %in% c("None", "Other")) %>%
    count(title, language, .drop = FALSE) %>% 
    complete(title, language) %>%
    replace_na(list(n = 0)) %>%
    group_by(title) %>%
    mutate(prop = prop.table(n))

p <- programming %>% 
    mutate(text = paste0("Language: ", language, "\n", 
                         "Job title: ", title, "\n", 
                         "Count: ", n, "\n",
                         "Proportion: ", round(prop, 3))) %>%
    ggplot(aes(language, title, fill=prop, text=text)) +
    geom_tile() +
    scale_fill_gradient(low="white", high="blue") +
    labs(
        title = "Users' favorite programming language",
        caption = glue("Author: celeritasML
                   Source: Kaggle")) +
    theme(axis.ticks.x = element_blank(),
          axis.text.x = element_text(angle=90, hjust=1),
          axis.title = element_blank(),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank())
ggplotly(p, tooltip="text")
```

```{r q7-ide}
ide <- job.dat %>% 
    select(c(Q5, starts_with("Q9_"))) %>%
    gather("fake_key", "IDE", -Q5, na.rm = T) %>%
    rename(title = Q5) %>%
    select(-fake_key) %>%
    mutate(IDE = case_when(
        IDE == "Visual Studio Code (VSCode)" ~ "VSCode",
        IDE == "Jupyter (JupyterLab, Jupyter Notebooks, etc)" ~ "Jupyter Notebook",
        TRUE ~ IDE
    )) %>%
    filter(!IDE %in% c("None", "Other")) %>%
    count(title, IDE, .drop = FALSE) %>% 
    complete(title, IDE) %>%
    replace_na(list(n = 0)) %>%
    group_by(title) %>%
    mutate(prop = prop.table(n))

p <- ide %>% 
    mutate(text = paste0("IDE: ", IDE, "\n", 
                         "Job title: ", title, "\n", 
                         "Count: ", n, "\n",
                         "Proportion: ", round(prop, 3))) %>%
    ggplot(aes(IDE, title, fill=prop, text=text)) +
    geom_tile() +
    scale_fill_gradient(low="white", high="blue") +
    labs(
        title = "Users' favorite IDE",
        caption = glue("Author: celeritasML
                   Source: Kaggle")) +
    theme(axis.ticks.x = element_blank(),
          axis.text.x = element_text(angle=90, hjust=1),
          axis.title = element_blank(),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank())
ggplotly(p, tooltip="text")
```

## Q8: Learning Sources

Where do they get and share the knowledge?

Survey questions Q39 (share and deploy), Q40 (learning resources), Q42 (Media sources).

```{r q8-learn}
learning_platform <- job.dat %>%
    select(c(Q5, starts_with("Q40_"))) %>%
    gather("fake_key", "learning", -Q5, na.rm = T) %>%
    rename(title = Q5) %>%
    select(-fake_key) %>%
    mutate(learning = case_when(
        learning == "Cloud-certification programs (direct from AWS, Azure, GCP, or similar)" ~ "Cloud-certif Programs",
        learning == "University Courses (resulting in a university degree)" ~ "University",
        TRUE ~ learning
    )) %>%
    filter(!learning %in% c("None", "Other")) %>%
    count(title, learning, .drop = FALSE) %>%
    complete(title, learning) %>%
    replace_na(list(n = 0)) %>%
    group_by(title) %>%
    mutate(prop = prop.table(n))

p <- learning_platform %>%
    mutate(text = paste0("Platform: ", learning, "\n",
                         "Job title: ", title, "\n",
                         "Count: ", n, "\n",
                         "Proportion: ", round(prop, 3))) %>%
    ggplot(aes(learning, title, fill=prop, text=text)) +
    geom_tile() +
    scale_fill_gradient(low="white", high="blue") +
    labs(
        title = "Users' favorite learning platforms",
        caption = glue("Author: celeritasML
                   Source: Kaggle")) +
    theme(axis.ticks.x = element_blank(),
          axis.text.x = element_text(angle=90, hjust=1),
          axis.title = element_blank(),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank())
ggplotly(p, tooltip="text")
```

```{r q8-share}
share_deploy <- job.dat %>% 
    select(c(Q5, starts_with("Q39_"))) %>%
    gather("fake_key", "share", -Q5, na.rm = T) %>%
    rename(title = Q5) %>%
    select(-fake_key) %>%
    mutate(share = case_when(
        share == "I do not share my work publicly" ~ "\'PRIVATE\'",
        TRUE ~ share
    )) %>%
    filter(!share %in% c("Other")) %>%
    count(title, share, .drop = FALSE) %>% 
    complete(title, share) %>%
    replace_na(list(n = 0)) %>%
    group_by(title) %>%
    mutate(prop = prop.table(n))

p <- share_deploy %>% 
    mutate(text = paste0("Platform: ", share, "\n", 
                         "Job title: ", title, "\n",
                         "Count: ", n, "\n",
                         "Proportion: ", round(prop, 3))) %>%
    ggplot(aes(share, title, fill=prop, text=text)) +
    geom_tile() +
    scale_fill_gradient(low="white", high="blue") +
    labs(
        title = "Users' favorite share platforms",
        x = "",
        y = "",
        caption = glue("Author: celeritasML
                   Source: Kaggle")) +
    theme(axis.ticks.x = element_blank(),
          axis.text.x = element_text(angle=90, hjust=1),
          axis.title = element_blank(),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank())
ggplotly(p, tooltip="text")
```

```{r q8-media}
media_source <- job.dat %>% 
    select(c(Q5, starts_with("Q42_"))) %>%
    gather("fake_key", "media", -Q5, na.rm = T) %>%
    rename(title = Q5) %>%
    select(-fake_key) %>%
    filter(!media %in% c("None", "Other")) %>%
    count(title, media, .drop = FALSE) %>% 
    complete(title, media) %>%
    replace_na(list(n = 0)) %>%
    group_by(title) %>%
    mutate(prop = prop.table(n)) %>%
    separate(media, into = c("media", "media_suffix"), sep = " \\(")

p <- media_source %>% 
    mutate(text = paste0("Platform: ", media, "\n", 
                         "Job title: ", title, "\n", 
                         "Count: ", n, "\n", 
                         "Proportion: ", round(prop, 3))) %>%
    ggplot(aes(media, title, fill=prop, text=text)) +
    geom_tile() +
    scale_fill_gradient(low="white", high="blue") +
    labs(
        title = "Users' favorite media source",
        caption = glue("Author: celeritasML
                   Source: Kaggle")) +
    theme(axis.ticks.x = element_blank(),
          axis.text.x = element_text(angle=90, hjust=1),
          axis.title = element_blank(),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank())
ggplotly(p, tooltip="text")
```


## Q9: Popular Packages

What data science packages do they use?

Survey questions Q14 (visualization libraries), Q16 (machine learning related libraries).

```{r q9-viz}
viz_lib <- dat %>%
  select(Q5, starts_with("Q14")) %>%
  select(-c(Q14_Part_11, Q14_OTHER))

viz_lib <- viz_lib %>%
  pivot_longer(cols=starts_with("Q14")) %>%
  select(-name) %>%
  drop_na() %>%
  filter(Q5!="Other")

ggplot(viz_lib) +
  geom_bar(aes(y = value, fill = value)) +
  facet_wrap(~ Q5) +
  scale_fill_brewer(palette = "Spectral") +
  labs(
    title = "Data science practitioners' favorite visualization libraries",
    x = "",
    y = "",
    caption = glue("Author: celeritasML
                   Source: Kaggle")
  ) +
  theme(
        axis.ticks.x = element_line(),
        axis.ticks.y = element_blank(),
        axis.text.x = element_text(size = 6),
        axis.text.y = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank())
```
```{r q9-ml}
ml_lib <- dat %>%
  select(Q5, starts_with("Q16")) %>%
  select(-c(Q16_Part_17, Q16_OTHER))

ml_lib <- ml_lib %>%
  pivot_longer(cols=starts_with("Q16")) %>%
  select(-name) %>%
  drop_na() %>%
  filter(Q5!="Other")

top_10_ml <- ml_lib %>%
  group_by(value) %>%
  summarize(count = n()) %>%
  arrange(desc(count)) %>%
  top_n(10)

ml_lib <- ml_lib %>%
  filter(value %in% top_10_ml$value)

ggplot(ml_lib) +
  geom_bar(aes(y = value, fill = value)) +
  facet_wrap(~ Q5) +
  scale_fill_brewer(palette = "Spectral") +
  labs(
    title = "Data science practitioners' favorite ML libraries",
    x = "",
    y = "",
    caption = glue("Author: celeritasML
                   Source: Kaggle")
  ) +
  theme(
        axis.ticks.x = element_line(),
        axis.ticks.y = element_blank(),
        axis.text.x = element_text(size = 6),
        axis.text.y = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank())
```

## Q10: Cloud Computing Platforms Preferences

Will a user's preference for cloud computing platforms affect his or her preference for other tools? For example, we want to know if an AWS EC2 dedicated user will actually prefer AWS S3 over other products.

Survey question Q29-A: computing products (Part_1)
Survey question Q30: Storage (Part_3, Part_4)
Survey question Q31-A: ML products (Part_1)


$$\chi^2=\sum\frac{(O_i-E_i)^2}{E_i}$$

```{r q10}
aws_user <- tibble(
  ec2 = dat$Q29_A_Part_1,
  s3 = dat$Q30_A_Part_3,
  efs = dat$Q30_A_Part_4,
  sagemaker = dat$Q31_A_Part_1,
  redshift = dat$Q32_A_Part_11,
  aurora = dat$Q32_A_Part_12,
  rds = dat$Q32_A_Part_13,
  dynamodb = dat$Q32_A_Part_14
  
) %>%
  mutate(ec2 = if_else(is.na(ec2), 0, 1),
         s3 = if_else(is.na(s3), 0, 1),
         efs = if_else(is.na(efs), 0, 1),
         sagemaker = if_else(is.na(sagemaker), 0, 1),
         redshift = if_else(is.na(redshift), 0, 1),
         aurora = if_else(is.na(aurora), 0, 1),
         rds = if_else(is.na(rds), 0, 1),
         dynamodb = if_else(is.na(dynamodb), 0, 1))
```

```{r q10-a}
chisq.test(aws_user$ec2, aws_user$s3)
```

```{r q10-b}
chisq.test(aws_user$ec2, aws_user$efs)
```

```{r q10-c}
chisq.test(aws_user$ec2, aws_user$sagemaker)
```



## Q11: AWS vs GCP

What is the overall AWS usage percentage among DS practitioners? Is it the same for Google Cloud?

$H_0: p_A=p_B$, $H_a:p_A\not=p_B$, where $A$ and $B$ can be replaced by AWS, Azure, or GCP, and $n_A$, $n_B$ are sample size of group $A$ and $B$ respectively.

The test statistic (z-statistic) can be calculated as follow:

$$z=\frac{p_A-p_B}{\sqrt{p(1-p)/n_A+p(1-p)/n_B}}$$

```{r q11}
cloud_comp <- tibble(
  aws_usage = dat$Q27_A_Part_1,
  azure_usage = dat$Q27_A_Part_2,
  gcp_usage = dat$Q27_A_Part_3
) %>%
  mutate(aws_usage = if_else(is.na(aws_usage), FALSE, TRUE),
         azure_usage = if_else(is.na(azure_usage), FALSE, TRUE),
         gcp_usage = if_else(is.na(gcp_usage), FALSE, TRUE))
```

```{r q11-a}
prop.test(c(sum(cloud_comp$aws_usage), sum(cloud_comp$azure_usage)),
          c(nrow(cloud_comp), nrow(cloud_comp)),
          alternative = "two.sided",
          correct = TRUE)
```

```{r q11-b}
prop.test(c(sum(cloud_comp$azure_usage), sum(cloud_comp$gcp_usage)),
          c(nrow(cloud_comp), nrow(cloud_comp)),
          alternative = "two.sided",
          correct = TRUE)
```

```{r q11-c}
prop.test(c(sum(cloud_comp$gcp_usage), sum(cloud_comp$aws_usage)),
          c(nrow(cloud_comp), nrow(cloud_comp)),
          alternative = "two.sided",
          correct = TRUE)
```

# Section 4: Conclusions

# Appendix: R scripts

```{r}
# R scripts here.
```

See footnote 1[^1].

[^1]: See GitHub repository [tufte](https://github.com/rstudio/tufte).

```{marginfigure}
Since the same codings will be used in Problem 1-c, I will just use this part as a setup.
```