---
title: "Data Science Career Paths & Skillsets in 2021"
subtitle: "A glimpse of Kaggle’s State of Data Science and Machine Learning Survey"
author:
  - Jingsong Gao, jg2109@georgetown.edu
  - Ercong Luo, el890@georgetown.edu
  - Rui Qiu, rq47@georgetown.edu
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_collapsed: true
    toc_depth: 3
    number_sections: true
    theme: lumen
    highlight: kate
---

```{r setup, include=FALSE}
# invalidate cache when the package version changes
knitr::opts_chunk$set(cache = TRUE,
                      echo = FALSE,
                      eval = TRUE,
                      tidy = FALSE,
                      warning = FALSE,
                      cache.extra = packageVersion('tidyverse'))
options(htmltools.dir.version = FALSE)

if (!require("pacman")) install.packages("pacman")
pacman::p_load(tidyverse, ggthemes, latex2exp, glue,
               hrbrthemes, plotly, stringr, DT, extrafont,
               tidymodels)

set.seed(511)

ggplot2::theme_set(
    theme_fivethirtyeight() +
    theme(
        text = element_text(family = "Roboto Condensed"),
        title = element_text(size = 14),
        plot.subtitle = element_text(size = 12),
        plot.caption = element_text(size = 10),
        axis.title = element_text(size = 14),
        axis.text = element_text(size = 12),
        panel.grid.minor.x = element_blank()
    )
)
```

```{marginfigure}
Since the same codings will be used in Problem 1-c, I will just use this part as a setup.
```

# Section 1: Introduction

Address the data science problem

# Section 2: Data Cleaning

```{r pkg, message=FALSE}
col_names <- names(read_csv(
  "data/kaggle_survey_2021_responses.csv",
  n_max=0))
dat <- read_csv(
  "data/kaggle_survey_2021_responses.csv",
  col_names = col_names, skip=2)

dat <- dat %>%
  filter(Q3=="United States of America" )

job.dat <- dat %>% 
    filter(Q5 %in% c("Data Analyst",
                     "Data Engineer",
                     "Data Scientist",
                     "Machine Learning Engineer",
                     "Software Engineer",
                     "Statistician",
                     "Student"))
```

# Section 3: The Analysis and Results

- List of questions.

```{marginfigure}
Since the one large data science problem can be divided into several smaller bits, we decide to combine the analysis and the interpretation sections together. And the questions will be addressed and discussed one by one.
```

```{marginfigure}
The following questions are about data science career paths.
```

## Q1: Job Titles Distributions

What percentage of the survey respondents are working under these job titles? 

```{r q1-job}
jtitle <- sort(table(dat$Q5), decreasing = T) %>% 
    as.data.frame() %>%
    as.tibble()
jtitle <- rename(jtitle, `Job Title` = Var1)

ggplot(jtitle, aes(x="", y=Freq, fill=`Job Title`)) +
    geom_bar(stat="identity", width=1, color="white") +
    coord_polar("y", start=0) + 
    theme_void() 
```

## Q2: Salaries

What are the statistics on salaries for these job titles?

The US minimum wage is 7.25 per hour, multiply that for a total $260$ typical number of workdays a year:

$$7.25 \frac{\$}{\text{hour}} \cdot 24 \frac{\text{hour}}{\text{day}} \cdot 260\frac{\text{day}}{\text{year}} = \$45240/\text{year}.$$

The new categories* for salaries will be:
- poverty: below the federal minimum wage
- low: 40,000 to 79,999
- medium: 80,000 to 124,999
- high: 125,000 to 199,999
- very high: 200,000 to 499,999
- highest: >= 500,000

*loosely based on US federal income tax brackets. 

```{r q2-salary}
# To change salary categories into FACTOR dtype with descending labels
poverty   <- c("$0-999", "1,000-1,999" , "2,000-2,999", "3,000-3,999", "4,000-4,999", "5,000-7,499", "7,500-9,999",
             "10,000-14,999", "15,000-19,999", "20,000-24,999", "25,000-29,999", "30,000-39,999") 
low       <- c("40,000-49,999", "50,000-59,999", "60,000-69,999", "70,000-79,999")
medium    <- c("80,000-89,999", "90,000-99,999", "100,000-124,999")
high      <- c("125,000-149,999", "150,000-199,999")
very_high <- c("200,000-249,999", "250,000-299,999", "300,000-499,999")
highest   <- c("$500,000-999,999", ">$1,000,000")

dat$Q25[dat$Q25 %in% poverty] <- "poverty"
dat$Q25[dat$Q25 %in% low] <- "low"
dat$Q25[dat$Q25 %in% medium] <- "medium"
dat$Q25[dat$Q25 %in% high] <- "high"
dat$Q25[dat$Q25 %in% very_high] <- "very high"
dat$Q25[dat$Q25 %in% highest] <- "highest"
dat$Q25 <- factor(dat$Q25, levels = c("poverty", "low", "medium", "high", "very high", "highest"), ordered = T)

data_side <- c("Data Scientist", "Data Analyst", "Business Analyst", "Data Engineer", "Statistician", "DBA/Database Engineer")
swe_side <- c("Software Engineer", "Machine Learning Engineer", "Program/Project Manager", "Product Manager") 
academia <- c("Student", "Other", "Research Scientist") 

dat[dat$Q5 %in% data_side & !is.na(dat$Q5) & !is.na(dat$Q25), ] %>%
  ggplot( aes(x=Q5, y=Q25, color = Q5)) +
    geom_count() + 
    ggtitle("Two-Way Salary Visualizations: Data-Oriented Jobs") +
    xlab("") +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))


dat[dat$Q5 %in% swe_side & !is.na(dat$Q5) & !is.na(dat$Q25), ] %>%
  ggplot( aes(x=Q5, y=Q25, color = Q5)) +
    geom_count() + 
    ggtitle("Two-Way Salary Visualizations: Engineering-Oriented Jobs") +
    xlab("") + 
    theme(axis.text.x = element_text(angle = 45, hjust = 1))

dat[dat$Q5 %in% academia & !is.na(dat$Q5) & !is.na(dat$Q25), ] %>%
  ggplot( aes(x=Q5, y=Q25, color = Q5)) +
    geom_count() + 
    ggtitle("Two-Way Salary Visualizations: Academic Jobs and Others") +
    xlab("") + 
    theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

## Q3: Requried Education Levels

What levels of education are required for these job titles?

## Q4: Gender Gap in Income

Is there a significant income gap between genders for these jobs?

## Q5: Skillset vs Income

What is the typical skill set for these jobs? How does it affect the pay rate?

```{r q5-skill}
skill.set <- job.dat %>% 
    select(c(Q5, starts_with("Q7_"), starts_with("Q9_"), 
             starts_with("Q12_"), starts_with("Q14_"),
             starts_with("Q16_"), starts_with("Q17_"),
             starts_with("Q18_"), starts_with("Q19_"))) %>%
    mutate(Total = "TotalHelper") %>%
    gather("fake_key", "skillset", -Q5, na.rm = T) %>%
    filter(!skillset %in% c("None", "Other")) %>%
    rename(title = Q5) %>%
    count(title, skillset) %>%
    group_by(title) %>%
    mutate(prop = round(n / max(n), 3)) %>%
    filter(prop >= 0.1 & skillset != "TotalHelper") %>%
    select(-n) %>%
    arrange(title, desc(prop))

datatable(skill.set, filter = 'top', width = 600)
```


## Q6: Correlation between Industry and Job

Is there a certain correlation between industry and the need for these jobs?

```{r q6-ind}
industry.dat <- job.dat %>%
    filter(Q5 != "Student") %>%
    select(Q5, Q20, Q25) %>%
    filter(Q20 %in% c("Academics/Education", 
                      "Accounting/Finance", 
                      "Computers/Technology",
                      "Insurance/Risk Assessment",
                      "Medical/Pharmaceutical",
                      "Online Service/Internet-based Services")) %>%
    mutate(Q25 = str_remove_all(Q25, "[$,]")) %>%
    mutate(Q25 = str_replace(Q25, ">1000000", "1000000-2000000")) %>%
    separate(Q25, into = c("salary_lb", "salary_ub"), sep = "-") %>%
    mutate(salary_lb = as.numeric(salary_lb)) %>%
    mutate(salary_ub = as.numeric(salary_ub))

p <- industry.dat %>% 
    count(Q5, Q20) %>%
    mutate(Q20 = fct_reorder(Q20, n, .fun="sum")) %>%
    rename(title=Q5, Industry=Q20, count=n) %>%
    ggplot(aes(x=title, y=count)) +
    geom_bar(stat = "identity") +
    coord_flip() +
    facet_wrap(~ Industry) +
    labs(
        title = "Users' work industry",
        caption = glue("Author: celeritasML
                   Source: Kaggle")) +
    theme(axis.ticks.x = element_blank(),
          axis.text.x = element_text(angle=90, hjust=1),
          axis.title = element_blank(),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank())
ggplotly(p)
```

```{r}
chisq.test(table(industry.dat$Q5, industry.dat$Q20))
```


```{r q6-ind-sal}
industry.dat %>% 
    mutate(Q20 = fct_reorder(Q20, salary_lb, .fun='length')) %>%
    ggplot(aes(x=Q20, y=salary_lb)) +
    geom_boxplot() +
    coord_flip() +
    facet_wrap(~ Q5) +
    labs(
        title = "Users' salary vs industry",
        caption = glue("Author: celeritasML
                   Source: Kaggle")) +
    theme(axis.ticks.x = element_blank(),
          axis.text.x = element_text(angle=90, hjust=1),
          axis.title = element_blank(),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank())
```


```{marginfigure}
The following questions are majorly about data science skills, tools and technologies.
```

## Q7: Languages and IDEs

What programming languages and IDEs do they use?

Survey questions Q7 (daily-used programming language), Q9 (IDE).

```{r q7-lan}
programming <- job.dat %>% 
    select(c(Q5, starts_with("Q7_"))) %>%
    gather("fake_key", "language", -Q5, na.rm = T) %>%
    rename(title = Q5) %>%
    select(-fake_key) %>%
    filter(!language %in% c("None", "Other")) %>%
    count(title, language, .drop = FALSE) %>% 
    complete(title, language) %>%
    replace_na(list(n = 0)) %>%
    group_by(title) %>%
    mutate(prop = prop.table(n))

p <- programming %>% 
    mutate(text = paste0("Language: ", language, "\n", 
                         "Job title: ", title, "\n", 
                         "Count: ", n, "\n",
                         "Proportion: ", round(prop, 3))) %>%
    ggplot(aes(language, title, fill=prop, text=text)) +
    geom_tile() +
    scale_fill_gradient(low="white", high="royalblue") +
    labs(
        title = "Users' favorite programming language",
        caption = glue("Author: celeritasML
                   Source: Kaggle")) +
    theme(axis.ticks.x = element_blank(),
          axis.text.x = element_text(angle=90, hjust=1),
          axis.title = element_blank(),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank())
ggplotly(p, tooltip="text")
```

```{r q7-ide}
ide <- job.dat %>% 
    select(c(Q5, starts_with("Q9_"))) %>%
    gather("fake_key", "IDE", -Q5, na.rm = T) %>%
    rename(title = Q5) %>%
    select(-fake_key) %>%
    mutate(IDE = case_when(
        IDE == "Visual Studio Code (VSCode)" ~ "VSCode",
        IDE == "Jupyter (JupyterLab, Jupyter Notebooks, etc)" ~ "Jupyter Notebook",
        TRUE ~ IDE
    )) %>%
    filter(!IDE %in% c("None", "Other")) %>%
    count(title, IDE, .drop = FALSE) %>% 
    complete(title, IDE) %>%
    replace_na(list(n = 0)) %>%
    group_by(title) %>%
    mutate(prop = prop.table(n))

p <- ide %>% 
    mutate(text = paste0("IDE: ", IDE, "\n", 
                         "Job title: ", title, "\n", 
                         "Count: ", n, "\n",
                         "Proportion: ", round(prop, 3))) %>%
    ggplot(aes(IDE, title, fill=prop, text=text)) +
    geom_tile() +
    scale_fill_gradient(low="white", high="royalblue") +
    labs(
        title = "Users' favorite IDE",
        caption = glue("Author: celeritasML
                   Source: Kaggle")) +
    theme(axis.ticks.x = element_blank(),
          axis.text.x = element_text(angle=90, hjust=1),
          axis.title = element_blank(),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank())
ggplotly(p, tooltip="text")
```

## Q8: Learning Sources

Where do they get and share the knowledge?

Survey questions Q39 (share and deploy), Q40 (learning resources), Q42 (Media sources).

```{r q8-learn}
learning_platform <- job.dat %>%
    select(c(Q5, starts_with("Q40_"))) %>%
    gather("fake_key", "learning", -Q5, na.rm = T) %>%
    rename(title = Q5) %>%
    select(-fake_key) %>%
    mutate(learning = case_when(
        learning == "Cloud-certification programs (direct from AWS, Azure, GCP, or similar)" ~ "Cloud-certif Programs",
        learning == "University Courses (resulting in a university degree)" ~ "University",
        TRUE ~ learning
    )) %>%
    filter(!learning %in% c("None", "Other")) %>%
    count(title, learning, .drop = FALSE) %>%
    complete(title, learning) %>%
    replace_na(list(n = 0)) %>%
    group_by(title) %>%
    mutate(prop = prop.table(n))

p <- learning_platform %>%
    mutate(text = paste0("Platform: ", learning, "\n",
                         "Job title: ", title, "\n",
                         "Count: ", n, "\n",
                         "Proportion: ", round(prop, 3))) %>%
    ggplot(aes(learning, title, fill=prop, text=text)) +
    geom_tile() +
    scale_fill_gradient(low="white", high="royalblue") +
    labs(
        title = "Users' favorite learning platforms",
        caption = glue("Author: celeritasML
                   Source: Kaggle")) +
    theme(axis.ticks.x = element_blank(),
          axis.text.x = element_text(angle=90, hjust=1),
          axis.title = element_blank(),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank())
ggplotly(p, tooltip="text")
```

```{r q8-share}
share_deploy <- job.dat %>% 
    select(c(Q5, starts_with("Q39_"))) %>%
    gather("fake_key", "share", -Q5, na.rm = T) %>%
    rename(title = Q5) %>%
    select(-fake_key) %>%
    mutate(share = case_when(
        share == "I do not share my work publicly" ~ "\'PRIVATE\'",
        TRUE ~ share
    )) %>%
    filter(!share %in% c("Other")) %>%
    count(title, share, .drop = FALSE) %>% 
    complete(title, share) %>%
    replace_na(list(n = 0)) %>%
    group_by(title) %>%
    mutate(prop = prop.table(n))

p <- share_deploy %>% 
    mutate(text = paste0("Platform: ", share, "\n", 
                         "Job title: ", title, "\n",
                         "Count: ", n, "\n",
                         "Proportion: ", round(prop, 3))) %>%
    ggplot(aes(share, title, fill=prop, text=text)) +
    geom_tile() +
    scale_fill_gradient(low="white", high="royalblue") +
    labs(
        title = "Users' favorite share platforms",
        x = "",
        y = "",
        caption = glue("Author: celeritasML
                   Source: Kaggle")) +
    theme(axis.ticks.x = element_blank(),
          axis.text.x = element_text(angle=90, hjust=1),
          axis.title = element_blank(),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank())
ggplotly(p, tooltip="text")
```

```{r q8-media}
media_source <- job.dat %>% 
    select(c(Q5, starts_with("Q42_"))) %>%
    gather("fake_key", "media", -Q5, na.rm = T) %>%
    rename(title = Q5) %>%
    select(-fake_key) %>%
    filter(!media %in% c("None", "Other")) %>%
    count(title, media, .drop = FALSE) %>% 
    complete(title, media) %>%
    replace_na(list(n = 0)) %>%
    group_by(title) %>%
    mutate(prop = prop.table(n)) %>%
    separate(media, into = c("media", "media_suffix"), sep = " \\(")

p <- media_source %>% 
    mutate(text = paste0("Platform: ", media, "\n", 
                         "Job title: ", title, "\n", 
                         "Count: ", n, "\n", 
                         "Proportion: ", round(prop, 3))) %>%
    ggplot(aes(media, title, fill=prop, text=text)) +
    geom_tile() +
    scale_fill_gradient(low="white", high="royalblue") +
    labs(
        title = "Users' favorite media source",
        caption = glue("Author: celeritasML
                   Source: Kaggle")) +
    theme(axis.ticks.x = element_blank(),
          axis.text.x = element_text(angle=90, hjust=1),
          axis.title = element_blank(),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank())
ggplotly(p, tooltip="text")
```


## Q9: Popular Packages

As for the popular packages, one can easily give some examples that no strangers to data science practitioners. But the real problem is, what can be learned from this? Some conditional probabilities will highlight the findings.

The data is a subset of survey answer to Q14 and Q16, which emphasize on visualization libraries and machine learning related libraries respectively. A set of niche job titles are excluded due to limited number of responses, such as "DBA/Database Engineer", "Currently not employed" and "Statistician." The remaining valid responses are visualized as barplots by group to show the count of responses that are familiar with those packages.

Additionally, as the number of machine learning libraries covered in the survey exceeds a reasonable number of meaning palette colors, only the top 10 libraries are kept in the visualization.

```{r q9-viz}
# Q9

viz_lib <- dat %>%
  select(Q5, starts_with("Q14")) %>%
  select(-c(Q14_Part_11, Q14_OTHER))

viz_lib <- viz_lib %>%
  pivot_longer(cols=starts_with("Q14")) %>%
  select(-name) %>%
  drop_na() %>%
  filter(!(Q5 %in% c("Other", "DBA/Database Engineer",
                   "Developer Relations/Advocacy",
                   "Currently not employed",
                   "Statistician", "Product Manager")))

ggplot(viz_lib) +
  geom_bar(aes(y = value, fill = value)) +
  facet_wrap(~ Q5) +
  scale_fill_brewer(palette = "Spectral") +
  labs(
    title = "Data science practitioners' favorite viz libraries",
    x = "",
    y = "",
    caption = glue("Author: celeritasML
                   Source: Kaggle")
  ) +
  theme(
        axis.ticks.x = element_line(),
        axis.ticks.y = element_blank(),
        axis.text.x = element_text(size = 6),
        axis.text.y = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank())
```
```{r q9-ml}
ml_lib <- dat %>%
  select(Q5, starts_with("Q16")) %>%
  select(-c(Q16_Part_17, Q16_OTHER))

ml_lib <- ml_lib %>%
  pivot_longer(cols=starts_with("Q16")) %>%
  select(-name) %>%
  drop_na() %>%
  filter(!(Q5 %in% c("Other", "DBA/Database Engineer",
                   "Developer Relations/Advocacy",
                   "Currently not employed",
                   "Statistician", "Product Manager")))

top_10_ml <- ml_lib %>%
  group_by(value) %>%
  summarize(count = n()) %>%
  arrange(desc(count)) %>%
  top_n(10)

ml_lib <- ml_lib %>%
  filter(value %in% top_10_ml$value)

ggplot(ml_lib) +
  geom_bar(aes(y = value, fill = value)) +
  facet_wrap(~ Q5) +
  scale_fill_brewer(palette = "Spectral") +
  labs(
    title = "Data science practitioners' favorite ML libraries",
    x = "",
    y = "",
    caption = glue("Author: celeritasML
                   Source: Kaggle")
  ) +
  theme(
        axis.ticks.x = element_line(),
        axis.ticks.y = element_blank(),
        axis.text.x = element_text(size = 6),
        axis.text.y = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank())
```

There are some obvious results worth mentioning. But before that, it is important to reiterate that the comparison is based on the calculation of some conditional probabilities according to Bayes' Theorem.

$$P(A|B) = \frac{P(B|A)\cdot P(A)}{P(B)} = \frac{P(A \cap  B)}{P(B)}$$

where $A$ and $B$ are events.

```{r}
viz_lib %>%
  group_by(Q5) %>%
  table() %>%
  addmargins() %>%
  print()

ml_lib %>%
  group_by(Q5) %>%
  table() %>%
  addmargins() %>%
  print()
```

- Both **Data Scientists** and **Students** are the two major respondents to package preferences/familiarity questions.

$$
\begin{split}
  P(\text{Student}|\text{Visualization}) &= \frac{739}{3923} \approx 0.1884\\
  P(\text{Data Scientist}|\text{Visualization}) &= \frac{1165}{3923} \approx 0.2970 \\
  P(\text{Student}|\text{Machine Learning}) &=\frac{732}{4038} \approx 0.1813\\
  P(\text{Data Scientist}|\text{Machine Learning}) &=\frac{1364}{4038} \approx 0.3378
\end{split}
$$

- Python-related packages or packages with Python support are overall more popular among the responses.
  - For example, `Altair`, `bokeh`, `geoplotlib`, `leaflet/folium`, `matplotlib`, `plotly/plotly express`, `seaborn` are categorized as Python-wise, while `D3.js`, `ggplot2`, `leaflet/folium`, `plotly/plotly express` and `shiny` are categorized as R-wise. Note that there are some overlapping among these packages since they do have multiple language support.
  - Similarly, `Huggingface`, `Keras`, `LightGBM`, `Prophet`, `PyTorch`, `scikit-learn`, `TensorFlow`, `Xgboost` are Python compatible. `Caret`, `Keras`, `LightGBM`, `Prophet`, `TensorFlow`, `tidymodels` and `Xgboost` and usable in R.

$$
\begin{split}
P(\text{Python}|\text{Visualization}) &= (33+115+96+80+1247+525+861)/3923 \approx 0.7538\\ 
P(\text{R}|\text{Visualization}) &= (132+622+80+525+212)/3923 \approx 0.4005\\
P(\text{Python}|\text{Machine Learning}) &= (132+549+187+93+527+1125+666+506)/4038 \approx 0.9373\\ 
P(\text{R}|\text{Machine Learning}) &= (148+549+187+93+666+105+506)/4038 \approx 0.5582\\
\end{split}
$$

There are more to be discovered depending on the perspectives.

## Q10: Cloud Computing Platforms Preferences

As the demand for computational power increases along with the amount of data involved in data science industry, cloud computing is a topic that any data science practitioners cannot avoid. The major

Specifically, will a user's preference for a cloud computing platforms affect his or her preferences for other tools?

Will a user's preference for cloud computing platforms affect his or her preference for other tools? For example, we want to know if an AWS EC2 dedicated user will actually prefer AWS S3 over other products.

Survey question Q29-A: computing products (Part_1)
Survey question Q30: Storage (Part_3, Part_4)
Survey question Q31-A: ML products (Part_1)


$$\chi^2=\sum\frac{(O_i-E_i)^2}{E_i}$$

```{r q10}
# Q10

aws_user <- tibble(
  ec2 = dat$Q29_A_Part_1,
  s3 = dat$Q30_A_Part_3,
  efs = dat$Q30_A_Part_4,
  sagemaker = dat$Q31_A_Part_1,
  redshift = dat$Q32_A_Part_11,
  aurora = dat$Q32_A_Part_12,
  rds = dat$Q32_A_Part_13,
  dynamodb = dat$Q32_A_Part_14
  
) %>%
  mutate(ec2 = if_else(is.na(ec2), 0, 1),
         s3 = if_else(is.na(s3), 0, 1),
         efs = if_else(is.na(efs), 0, 1),
         sagemaker = if_else(is.na(sagemaker), 0, 1),
         redshift = if_else(is.na(redshift), 0, 1),
         aurora = if_else(is.na(aurora), 0, 1),
         rds = if_else(is.na(rds), 0, 1),
         dynamodb = if_else(is.na(dynamodb), 0, 1))
```

```{r q10-a}
chisq.test(aws_user$ec2, aws_user$s3)
```

```{r q10-b}
chisq.test(aws_user$ec2, aws_user$efs)
```

```{r q10-c}
chisq.test(aws_user$ec2, aws_user$sagemaker)
```

- [ ] add a table or something else.

## Q11: AWS, Azure and GCP

What is the overall AWS usage percentage among DS practitioners? Is it the same for Google Cloud? What about Microsoft Azure?

$H_0: p_A=p_B$, $H_a:p_A\not=p_B$, where $A$ and $B$ can be replaced by AWS, Azure, or GCP, and $n_A$, $n_B$ are sample size of group $A$ and $B$ respectively.

The test statistic (z-statistic) can be calculated as follow:

$$z=\frac{p_A-p_B}{\sqrt{p(1-p)/n_A+p(1-p)/n_B}}$$

```{r q11}
# Q11

cloud_comp <- tibble(
  aws_usage = dat$Q27_A_Part_1,
  azure_usage = dat$Q27_A_Part_2,
  gcp_usage = dat$Q27_A_Part_3
) %>%
  mutate(aws_usage = if_else(is.na(aws_usage), FALSE, TRUE),
         azure_usage = if_else(is.na(azure_usage), FALSE, TRUE),
         gcp_usage = if_else(is.na(gcp_usage), FALSE, TRUE))
```

```{r q11-a}
prop.test(c(sum(cloud_comp$aws_usage), sum(cloud_comp$azure_usage)),
          c(nrow(cloud_comp), nrow(cloud_comp)),
          alternative = "two.sided",
          correct = TRUE)
```

```{r q11-b}
prop.test(c(sum(cloud_comp$azure_usage), sum(cloud_comp$gcp_usage)),
          c(nrow(cloud_comp), nrow(cloud_comp)),
          alternative = "two.sided",
          correct = TRUE)
```

```{r q11-c}
prop.test(c(sum(cloud_comp$gcp_usage), sum(cloud_comp$aws_usage)),
          c(nrow(cloud_comp), nrow(cloud_comp)),
          alternative = "two.sided",
          correct = TRUE)
```

- explain
- CI

## Q12 Linear Regression on Salary

Honestly speaking, the best way to approach this question is to apply a logistic regression on the the dependent variable `salary` due to its categorical nature. However, in order to experiment a linear model, we take a bold move to transform the `salary` back to continuous by randomly sampling values in between the pay levels. Three typical languages used by data science practitioners are included as well: Python, R and SQL.

```{r q12-lm}
# Q12

lm_dat <- dat %>%
  select(Q1, Q2, Q4, Q5, Q6,
         Q7_Part_1, Q7_Part_2, Q7_Part_3,
         Q25) %>%
  drop_na(Q25) %>%
  rename(age_group = Q1,
         gender = Q2,
         education = Q4,
         job = Q5,
         experience = Q6,
         salary = Q25,
         python = Q7_Part_1,
         r = Q7_Part_2,
         sql = Q7_Part_3)

set.seed(511)

# - poverty: below the federal minimum wage
# - low: 40,000 to 79,999
# - medium: 80,000 to 124,999
# - high: 125,000 to 199,999
# - very high: 200,000 to 499,999
# - highest: >= 500,000

lm_dat <- lm_dat %>%
  rowwise() %>%
  mutate(salary = case_when(
    salary == "poverty" ~ sample(1:39999, 1),
    salary == "low" ~ sample(40000:79999, 1),
    salary == "medium" ~ sample(80000:124999, 1),
    salary == "high" ~ sample(125000:199999, 1),
    salary == "very high" ~ sample(200000:499999, 1),
    salary == "highest" ~ sample(500000:2000000, 1)
  )) %>%
  ungroup()

ggplot(lm_dat, aes(x=salary)) +
  geom_histogram() +
  labs(
        title = "Histogram of salaries",
        caption = glue("Author: celeritasML
                   Source: Kaggle"))

ggplot(lm_dat, aes(x=log(salary))) +
  geom_histogram() +
  labs(
        title = "Histogram of log(salary)",
        caption = glue("Author: celeritasML
                   Source: Kaggle"))

lm_dat <- lm_dat %>%
  mutate(python = if_else(is.na(python), 0, 1),
         r = if_else(is.na(r), 0, 1),
         sql = if_else(is.na(sql), 0, 1))
```

```{r q12-lm-model}
model1 <- lm(log(salary) ~ . + sql:r + sql:python + r:python + python:r:sql +
               job:sql + job:r + job:python + job:experience +
               job:education, data=lm_dat)
MASS::stepAIC(model1)
```

Besides `summary()` table,

```{r q12-final-model}
model2 <- lm(log(salary) ~ age_group + gender + education + job + experience + 
    python + sql + python:sql, data = lm_dat)

anova(model2)
```


# Section 4: Conclusions

- findings as bullet points:

1. We don't have a one-sentence conclusion to summarize all the findings above. However, what we learned from this project is more about...
2. gender inequality
3. ...

- Besides the loose collection of findings above, we believe that (ML/cloud computing) is the general trend of data science in the next couple of years.

- More to discover from the survey data. Comparison with historical data blah blah blah.

- Limitation. (from our proposal, last paragraph). Kaggle users != ds practitioners; lots of students -> not representative enough to reflect the reality in the industry.




# Appendix: R scripts

```{r ref.label=knitr::all_labels(), echo=TRUE, eval=FALSE}
```

See footnote 1[^1].

[^1]: See GitHub repository [tufte](https://github.com/rstudio/tufte).
